---
title: "ASSIGNMENT"
author: "Takunda Nyenya"
date: "18/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#BASICS

##1.1 Variables in R

What is the value of " B " after running these command? 

```{r}
a <- 5.0
c <- 10
t <- 12
B <- a+c/t
```
The value of B is: `r round (B,2)`

What type of vector do you get in each of the following?
```{r}
vec_1 <- c(44, 5, 0, "k")
vec_2 <- c(1, 2, 3, FALSE)
vec_3 <- c("a", "b", "c", TRUE)
vec_4 <- c(1, 2, 3, "4")
```
vec_1 is a numerical vector.

vec_2 is a numerical vector.

vec_3 is a character vector.

vec_4 is a numerical vector.
 


##1.2 Conditional subsetting

Consider the following vector:
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)
```

##1.2.1. Create a vector vec_1 with no missing values using conditional subsetting and explain your code.
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)
vec[!is.na(vec)]#THIS CODE ONLY DISPLAYS ONLY THE VALUES IN THE VECTOR
```


## 1.2.2. Create a vector of indexes of missing values using conditional subsetting and explain your code.
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)
 vec[is.na(vec)]#THIS CODE WILL ONLY SHOW THE MISSING VALUES

```

## 1.2.3. Write a code to determine how many missing values are in vector.
```{r}
vec <- c(141.2741, 141.8392, 140.5730,NA, 141.3571, 139.8689, 
         137.9384, NA,139.6620, 138.2521, 138.8635, 141.3885, NA)
length(vec[is.na(vec)])
```

## 1.2.4. Write a R code to create a data frame which contains details of 3 students.
```{r}
name<- c("LUCY", "KUNASHE", "GRACE")
surname <- c("NYENYA", "NYENYA", "NYENYA")
age <- c(43 , 16 , 5)
gender<-c("FEMALE", "MALE", "FEMALE")
df <- data.frame(name,surname,age,gender); df
 
```
#2. FUNCTIONS IN R

## 2.1. Create a function which returns a mean of an input vector and explain your code.
```{r}
mean <- function(vector){
 z<-mean(vector,na.rm=TRUE)
 return(z)
  }
#example:
mean<-c(1,2,3)
mean(mean)
```
Explanation: I created a function which returns the mean of a vector

## 2.2. Create a function which take no inputs , it reads a Titanic .csv file and returns its content and explain your code.
```{r}
read.csv('train.csv') #THIS CODE READS THE CSV FILE
```
Explanation : The function reads a csv file without inputting anything


#3. Conditional and Loops

## 3.1. Write a function that take an input below and returns a transformed vector below into 1 when its element is male and 2 when its female using ifelse.
```{r}
g <- c("MALE","FEMALE","FEMALE","MALE","FEMALE","MALE","MALE","FEMALE", "MALE","MALE","FEMALE","MALE","MALE",
"FEMALE","MALE","MALE","MALE")

```
code:
```{r}
Gender<-function(){
  y<-ifelse(g == "MALE",1,2)
  return(y)
}
Gender()
```


##3.2. Generate x from normal distribution and sample with replacement "M" and "F" for male and female respectively by running the code below.
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)



```


## 3.2.1. Use the "cbind " function in R to create one dataset and assign it to df.
```{r}

x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
df <- cbind(x,y)
print (df)
```


##3.2.2. Convert df from 3.2.1 into a data.frame 
```{r}
x <- rnorm(1000, 120, 5)
y <- sample(c("M","F"),size = 1000, replace = TRUE)
df <- data.frame(df)
print (df)

```


## 3.2.3. Write a code using tidyverse to compute a number of males ("M") and females ("F") in 3.2.1 
```{r}
library(tidyverse)

df %>% group_by(y) %>% count(
  
)

```


## 3.2.4. Create a bar plot using ggplot to visualize your results in 3.2.3
```{r}
library(ggplot2)

df %>% group_by(y) %>% count() %>% 
ggplot(aes(x=y,y=n,))+
  geom_col()+
  labs(y="Total" , x="Gender", title="Bar Plot")
  

```


## 3.2.5. Use mutate function to add sum of x per "M" and "F" and plot the results using sum of x per "M" and "F".
```{r}
library(readxl)
df_new <-df %>% group_by(y) %>% count %>% 
  mutate(per_no=(n/1000*100)) 
ggplot(df_new, aes(y, y=per_no))+
  geom_col()


```

